Episodic memory: mental time travel or a quantum `memory wave' function?

### 2018-12-19 main.tex
added new figures to paper
working on results section

\section*{Shared contexts of temporally proximal and remote events}

\begin{figure}[tp] \centering \includegraphics[width=0.75\textwidth]{figs/pres_rec_corrmats} \caption{\textbf{Temporal correlation matrices.}  Each panel displays the timepoint-by-timepoint correlation matrix for presented stimuli (original experience, left panels) and participants' memories of those stimuli (recalled experience, right panels).  }
\label{fig:corrmats}
\end{figure}

>>>>

\section*{How do our experiences and memories unfold over time?}
Although real-world experiences can exhibit rich temporal structures, it can be illustrative to consider the transformation of how we experience much simpler and less naturalistic events (e.g. studying random word lists) versus how we later remember those experiences.  Figure~\ref{fig:corrmats}A displays an average temporal correlation matrix describing the similarity structure of 1072 random word lists, each comprising 16 words.  I applied a topic model (trained on a curated corpus of Wikipedia articles) to yield a 100-dimensional embedding of each presented word~\citep{BleiEtal03}.  The left panel displays the average correlations between the topic vectors of each pair of presented words as a function of their presentation positions.  As shown, nearly all of the correlation ``mass'' is along the diagonal, and the off-diagonal entries are very close to zero.  Because the word orders were randomized, the temporal correlation matrix reflects that there are no consistent similarity patterns between any given pair of timepoints.  The right panel of Figure~\ref{fig:corrmats}A displays the average temporal correlation matrices describing participants' \textit{recalls} of those same lists.  Although the recall matrix exhibits the strongest correlations along its diagonal, some correlation mass bleeds into nearby recalls, reflecting (in part) participants' tendancies to recall semantically related words successively.\footnote{Note that participants recalled different numbers of words from each studied list; I used linear interpolation to resample each recall correlation matrix that preserved the size of the presentation correlation matrix.  This process resulted in some additional blurring along the diagonal.  In other words, the blurring along the diagonal reflects two phenomena: participants' tendancies to semanticly cluster their recalls, and an artifact of the temporal resampling.}  Differences in the temporal correlation structure between the original lists of words and the orders of participants' recalls illustrate one way that our memory systems can reorganize our experiences such that we recall them in a different way than we originally experienced them.

\begin{figure}[tp] \centering \includegraphics[width=0.75\textwidth]{figs/pres_rec_corrmats} \caption{\textbf{Temporal correlation matrices.}  Each panel displays the timepoint-by-timepoint correlation matrix for presented stimuli (original experience, left panels) and participants' memories of those stimuli (recalled experience, right panels).  \textbf{A. Presentation and recall of random word lists.}  The panels reflect temporal correlation matrices between the topic vectors of presented words (left) and recalled words (right) over 1072 randomly ordered 16-word lists.  \textbf{B. Presentation and recall of categorized word lists.}  Same format as Panel A, but here participants have studied and recalled 480 categorized word lists, where the word presentations were ordered by category.  \textbf{C. Presentation and recall of a television episode.} Each moment of a television episode (left) and participants' retellings of what happened in the episode (right) are characterized by topic vectors that reflect the moments' semantic contents~\citep{HeusEtal18c}.}
\label{fig:corrmats}
\end{figure}

Whereas random word lists exhibit no stable off-diagonal correlations, one can induce temporal correlations in random word lists by manipulating the presentation orders of words on each list.  The left panel of Figure~\ref{fig:corrmats}B displays the average temporal correlation matrix exhibited by 240 16-word lists, each of whose items are drawn from 4 randomly selected categories~\citep{ZimaEtal18}.  The words were then presented to participants in categorized order, creating a block-diagonal correlation structure.  The right panel displays the temporal correlation matrix of the participants' recalls.  Despite

\begin{figure}[tp] \centering \includegraphics[width=\textwidth]{figs/reinstatement} \caption{\textbf{Mental time travel in random word lists and video.}  }
  % colors: individual participants
  % black: average across participants
  % gray: autocorrelation with topic vector at timepoint
  % circles: local maxima in the across-participants average (i.e., movie events that share structure with the given timepoint that is distinct from other temporally proximal events)
\label{fig:reinstatement}
\end{figure}
